Magnetised Polish doughnuts revisited

### 2017-07-03 paper_revised/reply_referee.tex

\bigskip

{\bf R:} - Sec. 2.1: they introduce the equation of state (13) but do not discuss
how they treat the specific internal energy that is part of the
enthalpy used in (13). A comment should be added.

\bigskip

{\bf A:} In our work we are following the same procedure as in Komissarov (2006), where no comment on the treatment
of the specific internal energy is explicitly provided. We have simply followed the same approach and have used the
same polytropic relationship between the enthalpy and the fluid pressure, Eq.~(12), and between the rest-mass density and the
fluid pressure (as can be read in the paragraph following Eq.~(27)). Thanks to the remark of the referee we now realise 
that by doing this we (and Komissarov) are making an implicit assumption about the internal energy. By looking at the 
expression of the stress-energy tensor in Eq.~(9) of the manuscript, we see that the fluid enthalpy $w$ includes the 
rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition 
of this quantity (in units with $c=1$) is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is the specific internal energy and 
$p$ is the fluid pressure. From a thermodynamical point of view a fluid is non-relativistic in the limit $h=1$. Therefore,
since both equations of state we use are polytropes (relating $p$ with $w$ or with $\rho$ in the same functional form) we 
conclude that we are implictly assuming that $h=1$ in our models, i.e.~the disks we build in our procedure are non-relativistic 
from a thermodynamical point of view.

We have included the following comment at the end of the paragraph following Eq.~(27): 
``We notae that the fluid enthalpy $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where
$h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use poytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implictly assuming that $h=1$, i.e.~the discs we build in our procedure are non-relativistic from a thermodynamical point of view."

To avoid confusion, we have changed the name of variable used for the step of the Simpson's rule from $h$ to $\Delta r$.

@@

\bigskip

{\bf R:} - Sec. 2.1: I am not clear how the magnetic field is actually
prescribed. The author mention that the only nonzero component of
the magnetic field in the comoving frame is the azimuthal one but I am
still unclear how the strength of this component is specified. A
comment should be added.

\bigskip

{\bf A:} We have not included further details about the prescription of the magnetic field because, in practice, in our procedure we 
only need to use $p_m\equiv b^2/2$,
and not the individual components of the magnetic-field themselves. We believe that including those extra details in the manuscript does not
provide relevant information. If interested, the referee is addressed to appendix A of Komissarov 
(2006) for the derivation of the explicit relations (which we do not employ anyway).

@@

\bigskip

{\bf R:} - Sec. 2.1: More on this topic: do the fluid isobaric surfaces (i.e. the
surfaces of constant fluid pressure) coincide with the magnetic
isobaric surfaces (i.e. the surfaces of constant magnetic pressure)?
How are these surfaces related to the iso-density surfaces? A comment
should be added.

\bigskip

{\bf A:} They do not coincide, as can be inferred from Eqs.~(12) and (14). Also notice that the values of $r_{\rm max}$ and $r_{\rm m,max}$ reported in the tables, do not coincide, which shows that fluid isobaric surfaces and magnetic isobaric surfaces cannot coincide. On the other hand, the iso-density surfaces and the fluid isobaric surfaces do coincide, since both variables are related by the barotropic equation of state.

We have included the following text at the end of the first paragraph of the results section: ``We note in particular that the radii of the maximum fluid pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, respectively, for the same value of $\beta_{\mathrm{m}_{\mathrm{c}}}$, never coincide. This also reflects the fact that constant fluid pressure surfaces do not coincide with constant magnetic pressure surfaces."

@@

\bigskip

{\bf R:} - Sec. 3: the authors discuss that the rest-mass density distribution can
``be trivially obtained inverting the barotropic relation, $p = K
\rho^{\kappa}$". I am not sure this is so trivial unless an assumption is
made on the specific internal energy; the authors should discuss more
clearly how they actually obtain the density from Eqs. (12) and (14).

\bigskip

{\bf A:} 
We have addressed this issue previously, while responding to a related query raised by the referee.

%The specific internal energy $\varepsilon$ is obtained by combining the ideal fluid (gamma-law)
%equation of state which, in the notation used in the paper we write $P=(\kappa-1)\rho\varepsilon$,
%and the barotropic equation of state. Solving for $\varepsilon$ we obtain
%\begin{eqnarray*}
%\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}\,.
%\end{eqnarray*}
%We have incorporated the following statements at the end of the paragraph after Eq.~(27):
%``We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
%this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
%$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$."

@@

\bigskip

{\bf R:} - Fig. 4: the authors do not discuss what is the largest magnetization
that they can handle. I agree that $\beta_{m_c} = 10^{-3}$ is probably
already quite high, but what happens for say $\beta_{m_c} = 10^{-6}$? Can
a solution still be found?

\bigskip

{\bf A:} We have built models with very small values of $\beta_{m_c}$, such as  $\beta_{m_c} = 10^{-20}$ (extremely high magnetisation) without encountering numerical difficulties. In practice we have found that there are no qualitative differences in the structure of the discs once $\beta_{m_c}$ becomes smaller than a sufficiently small value, about $\beta_{m_c}=10^{-3}$. This is
the reason why we choose that particular value as our lower limit in the figures and tables of the manuscript.

### 2017-07-04 paper_revised/paper_revised.tex
brief discussion at the end of appendix B

>>>>

\sg{It is relevant to note that the only reference to the explicit form of the metric it is done to show that $\mathcal{L} \neq 0$ and $\partial r \mathcal{L} \neq 0$ inside the disk. This means that our result holds for any stationary and axisymmetric metric if the aforementioned inequalities hold.

Also, if the angular momentum distribution allows the existence of a cusp and a center (and $r_{\mathrm{c}} \neq r_{\mathrm{cusp}}$) the result also holds for any angular momentum distribution.}
